Short_note_captured_phenotype

### 2025-01-30 Manuscript_data_analyses.qmd
Adding definitive code version

Strandings <- fread("~/ATER PAU 2024/Papier postdoc sensibilité bycatch/Strandings_journal_version.csv")

@@

Declaration_Bycatch = fread("~/ATER PAU 2024/Papier postdoc sensibilité bycatch/Declaration_Bycatch_journal_version.csv")

@@

OBSMER_Bycatch <- fread("~/ATER PAU 2024/Papier postdoc sensibilité bycatch/OBSMER_Bycatch_journal_version.csv")

@@

# paramètres à garder:

# Decl_Obs_Merged_Bycatch = all_merged[, c("ORIGIN", "ESP_LIB_FAO", "SECT_COD_SACROIS_NIV3", "SECT_COD_SACROIS_NIV3b", "TRIMESTRE", "AN", "YEAR_DCO", "YEAR_PHO", "ENGIN2", "ENGIN4", "MAILLAGE", "TARGET_TAXON", "TARGET_TAXON3", "TARGET_TAXON4", "poids_kg")]

@@

apply(Declaration_Bycatch[SPECIES %in% c("Delphinus delphis", "Phocoena phocoena"),], 2, function(x) table(is.na(x))) # in particular, targeted taxa were reported in ~68% of cases (very few missing values otherwise)

@@

observation_effort <- readRDS("~/ATER PAU 2024/Papier postdoc sensibilité bycatch/observation_effort.rds")

@@

allgeo <- readRDS("~/ATER PAU 2024/Papier postdoc sensibilité bycatch/allgeo_Journal_version.rds") # map of ICES division of interest

@@

ggsave("Figure_S2.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

ggsave("Figure_S3A.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

ggsave("Figure_S3B.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

ggsave("Figure_S3C.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

@@

@@

@@

ggsave("Figure_S5.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

@@

plot(ggemmeans(SPT_sensitivity_DEL_sex, terms = "TRIMESTER")) +
  ggtitle("Predicted probabilities of by-caught individual's sex\nper trimester") +
  ylab("By-caught individual's sex (probability)") +
  theme(text = element_text(size = 30))

@@

ggplot(dcast(na.omit(rbind(OBSMER_Bycatch[!is.na(SEX) & SPECIES %in% c("Delphinus delphis", "Phocoena phocoena"), .(SEX, SPECIES, "Captures")], Strandings[!is.na(SEX), .(SEX, SPECIES, "Strandings")])), SEX + SPECIES + V3 ~ ., fun.aggregate = length), aes( x =V3, y = .)) +
  geom_bar(aes(fill = SEX), stat = "identity", position = "dodge") + 
  facet_wrap(~ SPECIES, scales = "free_y") +
  ggtitle("Sex distribution of by-catches/strandings") + 
  labs(x = "Species and data sources", y = "Number of individuals", fill = "Sex") + 
  scale_fill_grey() +  theme(text = element_text(size = 30), strip.text = element_text(face = "italic")) +  theme(axis.text.x = element_text(angle = 55, hjust = 1))

@@

ggplot(rbind(Decl_Obs_Merged_Bycatch[SPECIES %in% c("Delphinus delphis", "Phocoena phocoena"), .(BODY_MASS, SPECIES, "Captures")], Strandings[!is.na(SEX), .(BODY_MASS, SPECIES, paste0("Strandings,\n", SEX))]), aes( x = V3, y = BODY_MASS)) +
  geom_violin() + geom_boxplot(width = 0.1) +  
  ggtitle("Body mass of captured/stranded individuals") + 
  labs(x = "Species, data sources and sex", y = "Body mass (kg)") + 
  theme(text = element_text(size = 30), strip.text = element_text(face = "italic")) +  theme(axis.text.x = element_text(angle = 55, hjust = 1))  +
  facet_wrap(~ SPECIES)

@@

ggplot(rbind(OBSMER_Bycatch[, .(paste0("*", SPECIES, "*", ", ", SEX, "<br>Captures"), BODY_LENGTH)], Strandings[, .(paste0("*", SPECIES, "*", ", ", SEX, "<br>Strandings"), BODY_LENGTH)]), aes( x = V1, y = BODY_LENGTH)) + 
  geom_violin(width = 1) + geom_boxplot(width = 0.1) + geom_point(data=biblio, mapping=aes(x = param, y = VALEUR, shape=ref), size=2) + 
  ggtitle("Body length of captured/stranded individuals\n(compared with literature life-history traits)") +
  labs(x = "Species, sex, data sources and literature estimates", y = "Body length (cm)", shape= "Bibliographic\nreferences") +  
  theme(text = element_text(size = 15))  +  theme(axis.text.x = element_text(angle = 60, hjust = 1))  + 
  scale_x_discrete(limits = c("*Delphinus delphis*, Female<br>Strandings","*Delphinus delphis*, Female<br>Captures","Sexual maturity<br>(females, *Delphinus delphis*)","Asymptotic length<br>(females, *Delphinus delphis*)","*Delphinus delphis*, Male<br>Strandings","*Delphinus delphis*, Male<br>Captures","Sexual maturity<br>(males, *Delphinus delphis*)","Asymptotic length<br>(males, *Delphinus delphis*)", "*Phocoena phocoena*, Female<br>Strandings","*Phocoena phocoena*, Female<br>Captures","Sexual maturity<br>(females, *Phocoena phocoena*)","Asymptotic length<br>(females, *Phocoena phocoena*)","*Phocoena phocoena*, Male<br>Strandings","*Phocoena phocoena*, Male<br>Captures","Sexual maturity<br>(males, *Phocoena phocoena*)","Asymptotic length<br>(males, *Phocoena phocoena*)")) +
    theme(axis.text = element_markdown())

@@

ggsave("Figure_S11.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

@@

@@

ggsave("Figure_S16A.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

 ggsave("Figure_S16B.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

# Susceptibility to be captured

@@

ggsave("Figure_S7A.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

ggsave("Figure_S7B.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

ggplot(Decl_Obs_Merged_Bycatch_explanatory[!is.na(TARGETED_TAXA) & SPECIES == "Delphinus delphis"], aes(x = forcats::fct_infreq(TARGETED_TAXA), fill = forcats::fct_infreq(TARGETED_ISSCAAP))) +
  geom_bar()  +
  labs(y = "Number of captures", x = "Targeted taxa", fill = "ISSCAAP groups") +
  ggtitle("Merged data - Common dolphin") +
  theme(text = element_text(size = 20))   +    theme(axis.text.x = element_text(angle = 65, hjust = 1)) +
  scale_fill_discrete(labels = c("31 - Pleuronectiformes (N=105)", "32 - Gadiformes (N=91)", "33 - Miscellaneous coastal fishes (N=68)", "34 - Miscellaneous demersal fishes  (N=17)",  "36 - Tunas, bonitos (N=17)", "37 - Miscellaneous pelagic fishes (N=12)", "57 - Squids, cuttlefishes, octopuses (N=5)", "35 - Herrings, sardines, anchovies  (N=3)", "43 - Lobsters, spiny-rock lobsters (N=3)",  "38 - Sharks, rays (N=2)"))

ggplot(Decl_Obs_Merged_Bycatch_explanatory[!is.na(TARGETED_TAXA) & SPECIES == "Phocoena phocoena"], aes(x = forcats::fct_infreq(TARGETED_TAXA), fill = forcats::fct_infreq(TARGETED_ISSCAAP))) +
  geom_bar()  +
  labs(y = "Number of captures", x = "Targeted taxa", fill = "ISSCAAP groups") +
  ggtitle("Merged data - Harbour porpoise") +
  theme(text = element_text(size = 20))   +    theme(axis.text.x = element_text(angle = 65, hjust = 1)) +
  scale_fill_discrete(labels = c("31 - Pleuronectiformes (N=44)", "34 - Miscellaneous demersal fishes  (N=32)", "32 - Gadiformes (N=12)", "33 - Miscellaneous coastal fishes (N=3)",  "38 - Sharks, rays (N=2)", "37 - Miscellaneous pelagic fishes (N=1)", "43 - Lobsters, spiny-rock lobsters (N=1)", "57 - Squids, cuttlefishes, octopuses (N=1)"))

@@

ggplot(data = Decl_Obs_Merged_Bycatch_BM_DCO, aes(BODY_MASS, MESH_SIZE, color=TARGETED_ISSCAAP)) + geom_jitter()
# For the interpretation of the latter effect: a large majority of large mesh size (>200mm) are associated with targeting miscellaneous demersal fishes (i.e. mainly mofishes or Zeidae)

@@

plot(ggemmeans(model_BM_DCO_sqrt, terms = "FISHING_GEAR_grouped"), show_data = T, jitter = c(0.7, 0), dot_alpha = 0.2) +
  ggtitle("Predicted by-caught individual's body mass\nas a function of used fishing gear") +
  labs(y = "Body mass (kg)", x = "Fishing gears")  +
  theme(text = element_text(size = 30)) 
ggsave("Figure_S14A.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

plot(ggemmeans(model_BM_DCO_sqrt, terms = "TARGETED_ISSCAAP"), show_data = T, jitter = c(0.7, 0), dot_alpha = 0.2) +
  ggtitle("Predicted by-caught individual's body mass\nas a function of targeted taxa") +
  labs(y = "Body mass (kg)", x = "Targeted taxa (ISSCAAP group)")  +
  theme(text = element_text(size = 30))   +  theme(axis.text.x = element_text(angle = 55, hjust = 1))
ggsave("Figure_S14B.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

ggsave("Figure_S14C.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

ggsave("Figure_S14D.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

ggsave("Figure_S14E.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

>>>>

Strandings <- fread("~/Strandings_journal_version.csv")

@@

Declaration_Bycatch = fread("~/Declaration_Bycatch_journal_version.csv")

@@

OBSMER_Bycatch <- fread("~/OBSMER_Bycatch_journal_version.csv")

@@

@@

apply(Declaration_Bycatch[SPECIES %in% c("Delphinus delphis", "Phocoena phocoena"),], 2, function(x) table(is.na(x))) # in particular, targeted taxa were reported in ~75% of cases (very few missing values otherwise)

@@

observation_effort <- readRDS("~/observation_effort.rds")

@@

allgeo <- readRDS("~/allgeo.rds") # spatial information file

allgeo=allgeo[allgeo$offname %in% c("27.4.c", "27.7.d", "27.7.e", "27.7.f", "27.7.g", "27.7.h", "27.7.j", "27.7.k", "27.8.a", "27.8.b", "27.8.c", "27.8.d"),] #selection of only ices division along the French coast

@@

ggsave("Figure_S3.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

@@

@@

@@

ggsave("Figure_S5A.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

ggsave("Figure_S5B.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

ggsave("Figure_S5C.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

@@

ggsave("Figure_S7.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

plot(ggemmeans(SPT_sensitivity_DEL_sex, terms = "TRIMESTER")) +
  ggtitle("Predicted probabilities of by-caught individual's sex\nper trimester") +
  ylab("By-caught individual's sex (probability)") +
  xlab("Trimester") +
  theme(text = element_text(size = 30))

@@

ggplot(dcast(na.omit(rbind(OBSMER_Bycatch[!is.na(SEX) & SPECIES %in% c("Delphinus delphis", "Phocoena phocoena"), .(SEX, SPECIES, "Declared/observed\nby-catches")], Strandings[!is.na(SEX), .(SEX, SPECIES, "Strandings")])), SEX + SPECIES + V3 ~ ., fun.aggregate = length), aes( x =V3, y = .)) +
  geom_bar(aes(fill = SEX), stat = "identity", position = "dodge") + 
  facet_wrap(~ SPECIES, scales = "free_y") +
  ggtitle("Sex distribution of by-catches/strandings") + 
  labs(x = "Species and data sources", y = "Number of individuals", fill = "Sex") + 
  scale_fill_grey() +  theme(text = element_text(size = 30), strip.text = element_text(face = "italic")) +  theme(axis.text.x = element_text(angle = 55, hjust = 1))

@@

ggplot(rbind(Decl_Obs_Merged_Bycatch[SPECIES %in% c("Delphinus delphis", "Phocoena phocoena"), .(BODY_MASS, SPECIES, "Declared/observed\nby-catches")], Strandings[!is.na(SEX), .(BODY_MASS, SPECIES, paste0("Strandings,\n", SEX))]), aes( x = V3, y = BODY_MASS)) +
  geom_violin() + geom_boxplot(width = 0.1) +  
  ggtitle("Body mass of by-caught/stranded individuals") + 
  labs(x = "Species, data sources and sex", y = "Body mass (kg)") + 
  theme(text = element_text(size = 30), strip.text = element_text(face = "italic")) +  theme(axis.text.x = element_text(angle = 55, hjust = 1))  +
  facet_wrap(~ SPECIES)

@@

ggplot(rbind(OBSMER_Bycatch[, .(paste0("*", SPECIES, "*", ", ", SEX, "<br>Declared/observed by-catches"), BODY_LENGTH)], Strandings[, .(paste0("*", SPECIES, "*", ", ", SEX, "<br>Strandings"), BODY_LENGTH)]), aes( x = V1, y = BODY_LENGTH)) + 
  geom_violin(width = 1) + geom_boxplot(width = 0.1) + geom_point(data=biblio, mapping=aes(x = param, y = VALEUR, shape=ref), size=2) + 
  ggtitle("Body length of by-caught/stranded individuals\n(compared with literature life-history traits)") +
  labs(x = "Species, sex, data sources and literature estimates", y = "Body length (cm)", shape= "Bibliographic references") +  
  theme(text = element_text(size = 15))  +  theme(axis.text.x = element_text(angle = 60, hjust = 1))  + 
  scale_x_discrete(limits = c("*Delphinus delphis*, Female<br>Strandings","*Delphinus delphis*, Female<br>Declared/observed by-catches","Sexual maturity<br>(females, *Delphinus delphis*)","Asymptotic length<br>(females, *Delphinus delphis*)","*Delphinus delphis*, Male<br>Strandings","*Delphinus delphis*, Male<br>Declared/observed by-catches","Sexual maturity<br>(males, *Delphinus delphis*)","Asymptotic length<br>(males, *Delphinus delphis*)", "*Phocoena phocoena*, Female<br>Strandings","*Phocoena phocoena*, Female<br>Declared/observed by-catches","Sexual maturity<br>(females, *Phocoena phocoena*)","Asymptotic length<br>(females, *Phocoena phocoena*)","*Phocoena phocoena*, Male<br>Strandings","*Phocoena phocoena*, Male<br>Declared/observed by-catches","Sexual maturity<br>(males, *Phocoena phocoena*)","Asymptotic length<br>(males, *Phocoena phocoena*)")) +
    theme(axis.text = element_markdown())

@@

@@

ggsave("Figure_S14A.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

ggsave("Figure_S14B.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

ggsave("Figure_S17A.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

 ggsave("Figure_S17B.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

# Vulnerability to by-cath

@@

@@

@@

ggplot(Decl_Obs_Merged_Bycatch_explanatory[!is.na(TARGETED_TAXA) & SPECIES == "Delphinus delphis"], aes(x = forcats::fct_infreq(TARGETED_TAXA), fill = forcats::fct_infreq(TARGETED_ISSCAAP))) +
  geom_bar()  +
  labs(y = "Number of captures", x = "Targeted taxa", fill = "ISSCAAP groups") +
  ggtitle("Merged data - Common dolphin") +
  theme(text = element_text(size = 20))   +    theme(axis.text.x = element_text(angle = 65, hjust = 1)) +
  scale_fill_discrete(labels = c("31 - Pleuronectiformes (N=109)", "32 - Gadiformes (N=92)", "33 - Miscellaneous coastal fishes (N=68)", "34 - Miscellaneous demersal fishes  (N=18)",  "36 - Tunas, bonitos (N=17)", "37 - Miscellaneous pelagic fishes (N=12)", "57 - Squids, cuttlefishes, octopuses (N=5)", "35 - Herrings, sardines, anchovies  (N=3)", "43 - Lobsters, spiny-rock lobsters (N=3)",  "38 - Sharks, rays (N=2)"))

ggplot(Decl_Obs_Merged_Bycatch_explanatory[!is.na(TARGETED_TAXA) & SPECIES == "Phocoena phocoena"], aes(x = forcats::fct_infreq(TARGETED_TAXA), fill = forcats::fct_infreq(TARGETED_ISSCAAP))) +
  geom_bar()  +
  labs(y = "Number of captures", x = "Targeted taxa", fill = "ISSCAAP groups") +
  ggtitle("Merged data - Harbour porpoise") +
  theme(text = element_text(size = 20))   +    theme(axis.text.x = element_text(angle = 65, hjust = 1)) +
  scale_fill_discrete(labels = c("31 - Pleuronectiformes (N=47)", "34 - Miscellaneous demersal fishes  (N=33)", "32 - Gadiformes (N=13)", "33 - Miscellaneous coastal fishes (N=3)",  "38 - Sharks, rays (N=2)", "37 - Miscellaneous pelagic fishes (N=1)", "43 - Lobsters, spiny-rock lobsters (N=1)", "57 - Squids, cuttlefishes, octopuses (N=1)"))

@@

ggplot(data = Decl_Obs_Merged_Bycatch_BM_DCO, aes(BODY_MASS, MESH_SIZE, color=TARGETED_ISSCAAP)) + geom_jitter()
# For the interpretation of the latter effect: a large majority of large mesh size (>200mm) are associated with targeting miscellaneous demersal fishes (i.e. mainly monkfishes or Zeidae)

@@

plot(ggemmeans(model_BM_DCO_sqrt, terms = "FISHING_GEAR_grouped"), show_data = T, jitter = c(0.7, 0), dot_alpha = 0.2) +
  ggtitle("Predicted by-caught individual's body mass\nas a function of used fishing gear") +
  labs(y = "Body mass (kg)", x = "Fishing gears")  +
  theme(text = element_text(size = 30)) 
ggsave("Figure_S11A.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

plot(ggemmeans(model_BM_DCO_sqrt, terms = "TARGETED_ISSCAAP"), show_data = T, jitter = c(0.7, 0), dot_alpha = 0.2) +
  ggtitle("Predicted by-caught individual's body mass\nas a function of targeted taxa") +
  labs(y = "Body mass (kg)", x = "Targeted taxa (ISSCAAP group)")  +
  theme(text = element_text(size = 30))   +  theme(axis.text.x = element_text(angle = 55, hjust = 1))
ggsave("Figure_S11B.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

ggsave("Figure_S11C.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

ggsave("Figure_S11D.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

ggsave("Figure_S11E.pdf", device = "pdf", units = "in", width = 14, height = 8.5)

@@

ggsave("Figure_S16.pdf", device = "pdf", units = "in", width = 14, height = 8.5)
